A Monte Carlo Synthetic Acceleration Method for the $SP_N$ Equations

### 2013-03-04 doc/spn/spn.tex
small updates to spn doc

%%---------------------------------------------------------------------------%%
\section{Introduction}
\label{sec:introduction}
The neutron transport problem is complicated. Solutions cover a large
phase space and the problems of interest are often geometry complex,
very large, or both requiring tremendous computational resources to
generate an adequate solution. Modern deterministic methods for large
scale problems are commonly variants on the discrete ordinates ($S_N$)
method \cite{evans_denovo:_2010}. For full reactor core neutronics
simulations, the $S_N$ method requires potentially trillions of
unknown angular flux moments to be computed to achieve good accuracy
for the repsonses to be computed. Other forms of the transport
problem, including the $P_N$ method, take on a simpler form than the
more common $S_N$ methods but lack in accuracy when compared while
still requiring considerable computational resources for solutions in
multiple dimensions.

@@

To further motivate moving in this direction, recent developments in
the Exnihilo neutronics package at Oak Ridge National Laboratory have
permitted generation of the $SP_N$ system of equations for detailed
full core reactor models. By fully forming these equations and
formulating them as a linear algebra problem, we now have access to
all of the modern advancements in computational linear algebra
including Krylov solvers for asymmetric systems and preconditioning
methods such as algebraic multigrid. This leads us to then explore the
applicability of our work in discrete Monte Carlo methods for linear
systems as a possible solution method for the $SP_N$ equations. If
formulated correctly, we hypothesis that signficant improvement in
usage of computational resources may be observed compared to modern
solution techniques such as those suggested. In addition, solving the
$SP_N$ equations in this way also breaks away from the $S_N$ forms of
parallelism where spatial parallelism is achieved by an efficient
parallel sweep, angular efficiency achieved by pipelining, and energy
parallelism achieved by decoupling the groups. With the $SP_N$
equations as a full matrix system, we now can parallelize the problem
as prescribed by the linear solver, which may be significantly more
scalable than current $S_N$ transport practices.

@@

%%---------------------------------------------------------------------------%%
\section{Derivation of the $SP_N$ Equations}
\label{sec:derivation}
In this section, we derive the $SP_N$ equations. We begin by stating
the general time-independent neutron transport equation followed by a
derivation of the $P_N$ equations in planar geometry for multiple
energy groups. From these equations, we then apply a set of
approximations to yield the multi-dimensional, multi-group $SP_N$
equations for fixed source problems.

@@

neutron transport equation for a single dimension for planar geometry:

@@

where $g$ denotes the group index of $0$ to $G$ groups where $G=N_g-1$
and the integration of the scattering emission term of energy has been

@@

or expanded:

>>>>

%%---------------------------------------------------------------------------%%
\section{Introduction}
\label{sec:introduction}
The neutron transport problem is complicated. Solutions cover a large
phase space and the problems of interest are often geometrically
complex, very large, or both, requiring tremendous computational
resources to generate an adequate solution. Modern deterministic
methods for large scale problems are commonly variants on the discrete
ordinates ($S_N$) method \cite{evans_denovo:_2010}. For full reactor
core neutronics simulations, the $S_N$ method requires potentially
trillions of unknown angular flux moments to be computed to achieve
good accuracy for the repsonses of interest. Other forms of the
transport problem, including the $P_N$ method, take on a simpler form
than the more common $S_N$ methods but lack in accuracy when compared
while still requiring considerable computational resources for
solutions in multiple dimensions.

@@

To further motivate moving in this direction, recent developments in
the Exnihilo neutronics package at Oak Ridge National Laboratory have
permitted generation of the $SP_N$ system of equations for detailed
full core reactor models. By fully forming these equations and
formulating them as a linear algebra problem, we now have access to
all of the modern advancements in computational linear algebra
including Krylov solvers for asymmetric systems and preconditioning
methods such as algebraic multigrid. This leads us to then explore the
applicability of our work in discrete Monte Carlo methods for linear
systems as a possible solution method for the $SP_N$ equations. If
formulated correctly, we hypothesize that signficant improvement in
usage of computational resources may be observed compared to modern
solution techniques such as those suggested. In addition, solving the
$SP_N$ equations in this way also breaks away from the $S_N$ forms of
parallelism where spatial parallelism is achieved by an efficient
parallel sweep, angular efficiency achieved by pipelining, and energy
parallelism achieved by decoupling the groups. With the $SP_N$
equations as a full matrix system, we now can parallelize the problem
as prescribed by the linear solver, which may be significantly more
scalable than current $S_N$ transport practices.

@@

%%---------------------------------------------------------------------------%%
\section{Derivation of the $SP_N$ Equations}
\label{sec:derivation}
In this section, we derive the $SP_N$ equations, closely following the
work of Evans. We begin by stating the general time-independent
neutron transport equation followed by a derivation of the $P_N$
equations in planar geometry for multiple energy groups. From these
equations, we then apply a set of approximations to yield the
multi-dimensional, multi-group $SP_N$ equations for fixed source
problems.

@@

neutron transport equation for a single dimension in planar geometry:

@@

where $g$ denotes the group index of $0$ to $G$ groups, $G=N_g-1$, and
the integration of the scattering emission term of energy has been

@@

and when expanded gives:
